ASTR480 Final Report:\\ \textit{Asteroids in TESS}

### 2024-09-04 FinalReport/FinalReportDraft.tex
interpolate at frames, not linspace
no one was at iso chat

\subsection{Ligthcurves and Periods}\label{SubSec:LCsandPs}

@@

%Periods and amplitueds for everything detected
The next part of my analysis has to do with determining the periods and amplitudes of each asteroid's light curve.
For this there are a few methods I can try; using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars or peel back a layer of abstraction and use the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\cite{Astropy2022}. 
Some trialling of both methods is needed, as preliminary testing reveals of interesting similarities and differences between the packages. 
There will be differences in the period between the matched points and the interpolated points, not just because of the difference in average flux, but also from the larger number of points. This difference needs to be carefully thought through to understand what is the more likely period.

>>>>

\subsection{Ligthcurves}\label{SubSec:Lightcurves}

@@

%TODO data cleaning explaination

\subsection*{Period detection}\label{SubSec:Periods}

%Periods and amplitueds for everything detected

The next part of my analysis has to do with determining the periods and amplitudes of each asteroid's light curve.

A widely used technique in determining the periods of astronomical data is the Lomb-Scargle periodogram \citep[\citet{Lomb1976,Scargle1982}, but see][for a review]{VanderPlas2018}.
This method can efficently take discrete timeseries data of which the observations are unevenly spaced and calculate the frequencies present in the sample.
TESS should have rather evenly spaced data, as the FFIs are produced on a set candance, but sometimes data will be missing.
For example it may have been sigma clipped due to a spike in brightness that is unphysical for an asteroid.
The Lomb-Scargle periodogram deals with this uneven sampling much better than either a fast fourier transform or a direct least squares fit.

There are a few different implementation of a Lomb-Scargle periodogram avalible in world of astronomy python packages.
I first tried using the \texttt{Lightkurve} \citep{Lightkurve2018} package built for period analysis of TESS (and Kepler) data of variable stars.
\texttt{Lightkurve} was easy to use but didn't allow for fine tuning of the periodogram, or give any model of the curve it produced.
Peeling back a layer of abstraction and useing the Lomb-Scargle periodogram as implemented by \texttt{Astropy}\citep{Astropy2022} was the next option.
There were interesting similarities and differences between the packages.
\texttt{Lightkurve} is based on the \texttt{Astropy} methods, but dose not give all of the functionallity, instead opting to simplify the process.
\texttt{Astropy} gives a statstic on how confident it is in the detected frequency is, the false alarm probability.
It also returns a model, in the form of

\begin{equation}
  \label{Eq:LCModel}
  F(t;f,\vec{\theta}\,) = \theta_0 + \theta_1\cdot\sin{2\pi ft} +\theta_2\cdot\cos{2\pi ft}
\end{equation}

where the flux, $F$, is given in terms of time, $t$, the calculated frequency, f, and the model parameter $\vec{\theta} = [\theta_0, \theta_1,\theta_2]$.
These parameters are useful in reconstructing the sinusoid without rerunning the periodogram.

There will be differences in the period between the matched points and the interpolated points, not just because of the difference in average flux, but also from the larger number of points. This difference needs to be carefully thought through to understand what is the more likely period.
